---
title: "Weekl 2"
output:
  pdf_document:
    toc: yes
    toc_depth: 6
    number_sections: yes
    latex_engine: xelatex
    df_print: kable
    highlight: tango
  html_notebook: default
  html_document:
    toc: yes
    toc_depth: '6'
    df_print: paged
  word_document:
    toc: yes
    toc_depth: '6'
---

# Week 2

## Reading from MySQL

### mySQL 

Free and widely used open source database software
Widely used in internet based applications
Data are structured in
- Databases
- Tables within databases
- Fields within tables
Each row is called a record

### Connecting and listing databases

```{r}
library(RMySQL)
ucscDb <-
  dbConnect(MySQL(), user = "genome", host = "genome-mysql.cse.ucsc.edu")
result <- dbGetQuery(ucscDb, "show databases;")

dbDisconnect(ucscDb)

```

```{r}
result
```

### Connecting to hg19 and listing tables

```{r}
hg19 <- dbConnect(MySQL(),user="genome", db="hg19",host="genome-mysql.cse.ucsc.edu")
allTables <- dbListTables(hg19)
length(allTables)
```

```{r}
allTables[1:5]
```

### Get dimensions of a specific table

```{r}
dbListFields(hg19,"affyU133Plus2")
```

```{r}
dbGetQuery(hg19, "select count(*) from affyU133Plus2")
```

### Read from the table

```{r,error=TRUE}
affyData <- dbReadTable(hg19, "affyU133Plus2")
head(affyData)
```

### Select a specific subset

```{r,error=TRUE}
query<- dbSendQuery(hg19, "select * from affyU133Plus2 where misMatches between 1 and 3")
affyMis <- fetch(query); quantile(affyMis$rnisMatches)
```
```{r,error=TRUE}
affyMisSmall <- fetch(query, n = lO)

dbClearResult(query)
```
```{r,error=TRUE}
dirn(affyMisSmall)
```

### Don't forget to close the connection

```{r}
dbDisconnect(hg19)
```

