---
title: "Weekl 2"
output:
  pdf_document:
    toc: yes
    toc_depth: 6
    number_sections: yes
    latex_engine: xelatex
    df_print: kable
    highlight: tango
  html_notebook: default
  html_document:
    toc: yes
    toc_depth: '6'
    df_print: paged
  word_document:
    toc: yes
    toc_depth: '6'
---

# Week 2

## Reading from MySQL

### mySQL 

Free and widely used open source database software
Widely used in internet based applications
Data are structured in
- Databases
- Tables within databases
- Fields within tables
Each row is called a record

### Connecting and listing databases

```{r}
library(RMySQL)
ucscDb <-
  dbConnect(MySQL(), user = "genome", host = "genome-mysql.cse.ucsc.edu")
result <- dbGetQuery(ucscDb, "show databases;")

dbDisconnect(ucscDb)

```

```{r}
result
```

### Connecting to hg19 and listing tables

```{r}
hg19 <- dbConnect(MySQL(),user="genome", db="hg19",host="genome-mysql.cse.ucsc.edu")
allTables <- dbListTables(hg19)
length(allTables)
```

```{r}
allTables[1:5]
```

### Get dimensions of a specific table

```{r}
dbListFields(hg19,"affyU133Plus2")
```

```{r}
dbGetQuery(hg19, "select count(*) from affyU133Plus2")
```

### Read from the table

```{r,error=TRUE}
affyData <- dbReadTable(hg19, "affyU133Plus2")
head(affyData)
```

### Select a specific subset

```{r,error=TRUE}
query<- dbSendQuery(hg19, "select * from affyU133Plus2 where misMatches between 1 and 3")
affyMis <- fetch(query); quantile(affyMis$rnisMatches)
```
```{r,error=TRUE}
affyMisSmall <- fetch(query, n = lO)

dbClearResult(query)
```
```{r,error=TRUE}
dirn(affyMisSmall)
```

### Don't forget to close the connection

```{r}
dbDisconnect(hg19)
```

## Reading from HDF5

### HDFS

Used for storing large data sets
Supports storing a range of data types
Heirarchical data format
groups containing zero or more data sets and metadata
- Have a group header with group name and list of attributes
- Have a group symbol table with a list of objects in group
datasets multidmensional array of data elements with metadata
- Have a headerwith name, datatype, dataspace, and storage layout
- Have a data array with the data
lhttp://www.hdfgroup.org/


### R HDF5 package

```{r, error=TRUE}
library(rhdf5)
created <- h5createFile("example.h5")
created
```

### Create groups

```{r, error=TRUE}
created <- h5createGroup("example.h5", "foo")
created <- h5createGroup("example.h5", "baa")
created <- h5createGroup("example.h5", "foo/foobaa")
h5ls("example.h5")
```

### Write to groups

```{r, error=TRUE}
A <- matrix(1:10,nr=5,nc=2)
h5write(A,"example.h5","foo/A")
B <- array(seq(0.1,2.0,by=0.1),dim = c(5,2,2))
attr(B,"scale") <- "liter"
```

```{r, error=TRUE}
h5write(B,"example.h5","foo/foobaa/B")
```


```{r, error=TRUE}
h5ls("example.h5")
```


### Write a data set

```{r}
df <- data.frame(1L:5L, seq(0,1,length.out=5),c("ab","cde","fghi","a","s"),stringsAsFactors = FALSE)
```

```{r, error=TRUE}
h5write(df,"example.h5","df")
```

```{r}
h5ls("example.h5")
```


### Reading data

```{r}
readA <- h5read("example.h5","foo/A")
readB <- h5read("example.h5","foo/foobaa/B")
readdf <- h5read("example.h5","df")
```

```{r}
readA
```

```{r}
readB
```

```{r}
readdf
```


### Writing and reading chunks

```{r, error=TRUE}
h5write(c(12,13,14),"example.h5","foo/A",index=list(1:3,1))
```


```{r}
h5read("example.h5","foo/A")
```


## Reading from The Web

### Webscraping

Webscraping: Programatically extracting data from the HTML code of websites.
It can be a great way to get data How Netflix reverse engineered Hollywood
Many websites have information you may want to programaticaly read
In some cases this is against the terms of service for the website
Attempting to read too many pages too quickly can get your IP address blocked

### Getting data off webpages - readlinesO

```{r,error=TRUE}
con <- url("https://scholar.google.com/citations?user=HI-I6C0AAAAJ&hl=en")
htmlCode <- readLines(con)
close(con)
htmlCode
```

### Parsing with XML

```{r}
library(XML)
```

```{r}
html <- htmlTreeParse(htmlCode, useInternalNodes = T)

xpathSApply(html,"//title",xmlValue)
```

```{r}
xpathSApply(html,"//td[@id='col-citedby']",xmlValue)
```

### GET from the httr package

```{r}
library(httr)
```

```{r,error=TRUE}
html2 = GET(url)
content2 <- content(html2,as="text")
parsedHTML <- htmlParse(content2,asText = TRUE)
xpathSApply(parsedHTML,"//title",xmlValue)
```

### Accessing websites with passwords

```{r}
pg1 = GET("http://httpbin.org/basic-auth/user/passwd")
pg1
```


```{r,error=TRUE}
pg2 = GET("http://httpbin.org/basic-auth/user/passwd",
authenticate( "user", "passwd"))
pg2
```


### Using handles 

```{r}
google = handle("http://google.com")
pgl = GET(handle=google,path="/")
pg2 = GET(handle=google,path="search")
```


## Reading From APIs

In general look at the documentation
httr allows GET, POST, PUT, DELETE requests if you are authorized
You can authenticate with a user name or a password
Most modern APls use something like oauth
httr works well with Facebook, Google, Twitter, Githb, etc.

Authenticate -> Connect -> retrive -> convert JSON -> consume 

## Reading From Other Sources 

### There is a package for that

Roger has a nice video on how there are R packages for most things that you will want to access.
Here l1m going to briefly review a few useful packages
In general the best way to find out if the R package exists is to Google 11data storage mechanism R
package11
- For example: 11MySQL R package11

### Interactingm ore directlyw ith files

file - open a connection to a text file
url - open a connection to a url
gzfile - open a connection to a .gz file
bzfile - open a connection to a .bz2 file
?connections for more information
Remember to close connections

### foreign package

Loads data from Minitab, S, SAS, SPSS, Stata,Systat
Basic functions read.too
- read.arff (Weka)
- read.dta (Stata)
- read.mtp (Minitab)
- read.octave (Octave)
- read.spss (SPSS)
- read.xport (SAS)
See the help page for more details lhttp://cran.r-project.org/web/packages/foreign/foreign.pdf

### Examples of other database packages

RPostresSQL provides a DBI-compliant database
Tutorial-https://code.google.com/p/rpostgre~, help
project.org/web/packages/RPostgreSQURPostgreSQL.pdf
connection from R.
file-[nttp://cra[B
RODBC provides interfaces to multiple databases including PostgreQL, MySQL, Microsoft Access
and SQLite. Tutorial - ffi!p://cran.r-project.org/web/packages/RODBC/vignettes/RODBC.pdf, help
file - http://cran.r-project.org/web/packages/RODBC/RODBC.pcflj
RM on go lhttp://cran.r-project.org/web/packages/RMongo/RMongo.pd~ (example of Rmongo
http://www.r-bloggers.com/r-and-mongodb/) and rmongodbl provide interfaces to MongoDb.


### Reading images

jpeg - ffi!p://cran.r-project.org/web/packages/jpeg/index.html
readbitmap - lhttp://cran.r-project.org/web/packages/readbitmap/index.html
png - http://cran.r-project.org/web/packages/png/i ndex. htm I
EBlmage (Bioconductor) - http://www.bioconductor.org/packages/2.13/bioc/html/EBlmage.htmll

### Reading GIS data

rd gal - http://cran.r-project.org/web/packages/rgdal/index. hTm]
rgeos - lhttp://cran.r-project.org/web/packages/rgeos/index.htmll
raster - [nttp://cran.r-project.org/web/packages/raster/index.html

### Reading music data

tu neR - lfittp: // cran.r -project.org/web/packages/tun ~
seewave - http://rug.mnhn.fr/seewave


