---
title: "R Notebook"
output: html_notebook
---
```{r}
load("data/samsungData.rda")
names(samsungData[1:12])
```
```{r}
table(samsungData$activity)
```
```{r}
par(mfrow = c(1, 2), mar = c(5, 4, 1, 1)) # set up the plotting area into 1 row and 2 columns, and set the margins to 5, 4, 1, 1 (bottom, left, top, right)
samsungData <- transform(samsungData, activity = factor(activity)) # convert the activity variable into a factor
sub1 <- subset(samsungData, subject == 1) # create a subset of the data with only subject 1
plot(sub1[, 1], col = sub1$activity, ylab = names(sub1)[1]) # plot the first variable in the subset, with the points colored by activity
plot(sub1[, 2], col = sub1$activity, ylab = names(sub1)[2]) # plot the second variable in the subset, with the points colored by activity
legend("bottomright", legend = unique(sub1$activity), col = unique(sub1$activity), pch = 1) # add a legend to the bottom right of the plot
```


The provided line of code in R is used to create a subset of the 'samsungData' dataframe.

The subset function is used to extract subsets of a dataframe that meet a particular condition. In this case, the condition is 'subject == 1'. This means it's taking only those rows from the 'samsungData' dataframe where the 'subject' column has a value equal to 1.


```{r}
myplclust <- function(hclust, lab = hclust$labels, lab.col = rep(1, length(hclust$labels)), hang = 0.1, ...) {
  ## modifiction of plclust for plotting hclust objects *in colour*!
  ## Copyright Eva KF Chan 2009
  ## Arguments:
  ##    hclust:    hclust object
  ##    lab:        a character vector of labels of the leaves of the tree
  ##    lab.col:    colour for the labels; NA=default device foreground colour
  ##    hang:     as in hclust & plclust
  ## Side effect:
  ##    A display of hierarchical cluster with coloured leaf labels.
  y <- rep(hclust$height, 2)
  x <- as.numeric(hclust$merge)
  y <- y[which(x < 0)]
  x <- x[which(x < 0)]
  x <- abs(x)
  y <- y[order(x)]
  x <- x[order(x)]
  plot(hclust, labels = FALSE, hang = hang, ...)
  text(x = x, y = y[hclust$order] - (max(hclust$height) * hang), labels = lab[hclust$order], col = lab.col[hclust$order], srt = 90, adj = c(1, 0.5), xpd = NA, ...) }
```
```{r}
distanceMatrix <- dist(sub1[, 1:3])
hclustering <- hclust(distanceMatrix)
myplclust(hclustering, lab.col = unclass(sub1$activity))
```



The unclass() function in R is used to remove the class attribute from an object. This function is often used when we want to convert a data object from one type to another.

If sub1$activity is a factor, for example, unclass(sub1$activity) will return an integer vector which is the internal representation of the factor.
```{r}
par(mfrow = c(1,2))
plot(sub1[,10], col = sub1$activity, ylab = names(sub1)[10], pch = 19)
plot(sub1[,11], col = sub1$activity, ylab = names(sub1)[11], pch = 19)
legend("bottomright", legend = unique(sub1$activity), col = unique(sub1$activity), pch = 1) # add a legend to the bottom right of the plot
```
```{r}
distanceMatrix <- dist(sub1[, 10:12])
hclustering <- hclust(distanceMatrix)
myplclust(hclustering, lab.col = unclass(sub1$activity))
```

```{r}
svd1 <- svd(scale(sub1[, -c(562, 563)])) # perform SVD on the scaled data, excluding the last two columns
par(mfrow = c(1, 2)) # set up the plotting area into 1 row and 2 columns
plot(svd1$u[, 1], col = sub1$activity, pch = 19)# plot the first column of U, with the points colored by activity
plot(svd1$u[, 2], col = sub1$activity, pch = 19)# plot the first column of U, with the points colored by activity
legend("bottomright", legend = unique(sub1$activity), col = unique(sub1$activity), pch = 1) # add a legend to the bottom right of the plot
```
```{r}
plot(svd1$v[,2], pch=19) # plot the second column of V
```
```{r}
names(svd1)
```
```{r}
maxContrib <- which.max(svd1$v[,2]) # find the index of the maximum value in the second column of V
distanceMatrix <- dist(sub1[, c(10,12, maxContrib)]) # create a distance matrix using the three variables
hclustering <- hclust(distanceMatrix)
myplclust(hclustering, lab.col = unclass(sub1$activity))
```
```{r}
names(samsungData)[maxContrib]
```
```{r}
kClust <- kmeans(sub1[,-c(562,563)], centers = 6)
table(kClust$cluster, sub1$activity)
```
```{r}
kClust <- kmeans(sub1[,-c(562,563)], centers = 6, nstart = 100)
table(kClust$cluster, sub1$activity)
```
```{r}
plot(kClust$center[1,1:10], pch = 19, ylab = "Cluster center", xlab = "")
```
