---
title: "R Notebook"
output: html_notebook
---

## Creating Vectors

The c() function can be used to create vectors of objects.

```{r}
x <- c(0.5,0.6) # numeric
x <- c(TRUE,FALSE) # logical 
x <- c(T,F) # logical 
x <- c("a","b","c") # character 
x <- 9:29 # integer (range)
x <- c(1+0i,2+4i) # complex
```

```{r}
x <- vector("numeric", length = 10)
x
```

## Mixing Objects

When different objects are mixed in a vector, coercion occurs so that every element in the vector is of the same class.

```{r}
y <- c(1.7, "a") ## character
y
```

```{r}
y <- c(TRUE, 2) ## numeric
y
```
```{r}
y <- c("a", TRUE) ## character
y
```

## Explicit Coercion

Objects can be explicitly coerced from one class to another using the as.* functions, if available.

```{r}
x<-0:6
class(x)
```
```{r}
as.numeric(x)
```
```{r}
as.logical(x)
```
```{r}
as.character(x)
```

Nonsensical coercion results in NAs.

```{r}
x <- c("a", "b", "c")
as.numeric(x)
```
```{r}
as.logical(x)
```

```{r}
as.complex(x)
```

## Matrices

Matrices are vectors with a dimension attribute. The dimension attribute is itself an integer vector of length 2 (nrow, ncol)

```{r}
m <- matrix(nrow = 2, ncol = 3)
m
```

```{r}
dim(m)
```
```{r}
attributes(m)
```

## Matrices (cont’d)

Matrices are constructed column-wise, so entries can be thought of starting in the “upper left” corner and running down the columns.

```{r}
m <- matrix(1:6, nrow = 2, ncol = 3)
m
```

Matrices can also be created directly from vectors by adding a dimension attribute.

```{r}
m<-1:10
m
```
```{r}
dim(m) <- c(2, 5)
m
```

## cbind-ing and rbind-ing

Matrices can be created by column-binding or row-binding with cbind() and rbind().

```{r}
x<-1:3
y<-10:12
```

```{r}
x
```
```{r}
y
```
```{r}
cbind(x,y)
```

```{r}
rbind(x,y)
```

## Lists

Lists are a special type of vector that can contain elements of different classes. Lists are a very important data type in R and you should get to know them well.

```{r}
x <- list(1, "a", TRUE, 1 + 4i)
x
```

