---
title: "Creating New Variables"
output:
  pdf_document:
    toc: yes
    toc_depth: 6
    number_sections: yes
    latex_engine: xelatex
    df_print: kable
    highlight: tango
  html_notebook: default
  html_document:
    toc: yes
    toc_depth: '6'
    df_print: paged
  word_document:
    toc: yes
    toc_depth: '6'
---

# Creating New Variables

## Why create new variables ?

Often the raw data won't have a value you are looking for
You will need to transform the data to get the values you would like
Usually you will add those values to the data frames you are working with
Common variables to create
- Missingness indicators
- "Cutting up" quantitative variables
- Applying transforms

## Creating sequences

Sometimes you need an index for your data set

```{r}
s1 <- seq(1,10,by=2)
s1
```

```{r}
s2 <- seq(1,10,length=3)
s2
```

```{r}
x <- c(1,3,8,25,100); 
seq(along = x)
```
When you use the along argument in the seq() function, it generates a sequence from 1 to the length of the object you have specified.


## Subsetting variables

```{r,error=TRUE}
rest <- read.csv("./data/rest.csv")
```


```{r}
rest$nearme <- rest$nghbrhd %in% c("Roland Park","Homeland")
```

```{r}
table(rest$nearme)
```

## Creating binary variables

```{r}
rest$zipWrong <- ifelse(rest$zipcode < 0, TRUE, FALSE)
table(rest$zipWrong, rest$zipcode < 0)
```

## Creatingc ategoricavl ariables

```{r}
rest$zipcode <- as.numeric(rest$zipcode)
```


```{r, error=TRUE}
rest$zipGroups <- cut(rest$zipcode, breaks = quantile(rest$zipcode))
table(rest$zipGroups)
```

## Easier cutting

```{r}
library(Hmisc)
```

```{r,error=TRUE}
rest$zipGroups <- cut2(rest$zipcode, g=4)
```

## Creating factor variables

```{r,error=TRUE}
rest$zcf <- factor(rest$zipcode)
rest$zcf[1:10]
```

```{r}
class(rest$zcf)
```

## Levels of factor variables

```{r}
yesno <- sample(c("yes", "no"), size = 10, replace = TRUE)
yesnofac <- factor(yesno, levels = c("yes","no"))
```

```{r}
yesno
```


```{r}
yesnofac
```

```{r}
relevel(yesnofac, ref = "yes")
```

```{r}
as.numeric(yesnofac)
```

## Using the mutate function

```{r}
library(Hmisc)
library(plyr)
rest2 <- mutate(rest, zipGroups=cut2(zipcode, g=4))
```


```{r}
table(rest2$zipGroups)
```

The mutate() function is a part of the dplyr package (and also available in plyr) in R. It's used to create new variables or modify existing ones in a data frame.

```{r}
str(mutate)
```

## Common transforms

`abs` ( x) absolute value
`sqrt` ( x) square root
`ceiling` ( x) ceiling(3.475) is 4
`floor` ( x) floor(3.475) is 3
`round` ( x, digi ts=n) roun(3.475,digits=2) is 3.48
`signif` ( x, digi ts=n) signif(3.475,digits=2) is 3.5
`cos(x)`, sin(x) etc.
`log ( x)` natural logarithm
`log2 ( x)`, logl o ( x) other common logs
`exp ( x)` exponentiating x

