---
title: "Managing Data Frames with dplyr - Basic Tools"
output:
  pdf_document:
    toc: yes
    toc_depth: 6
    number_sections: yes
    latex_engine: xelatex
    df_print: kable
    highlight: tango
  html_notebook: default
  html_document:
    toc: yes
    toc_depth: '6'
    df_print: paged
  word_document:
    toc: yes
    toc_depth: '6'
---

# Managing Data Frames with dplyr - Basic Tools

## Load the dplyr package

This step is important!

```{r}
library(dplyr)
```

## select

```{r}
chicago <- readRDS("data/chicago.rds")
dim(chicago)
```
```{r}
head(select(chicago, 1:5))
```


```{r}
names(chicago)[1:3]
```

```{r}
head(select(chicago, city:dptp))
```


In dplyr you can do

```{r}
head(select(chicago, -(city:dptp)))
```

Equivalent base R

```{r}
i <- match("city", names(chicago))
j <- match("dptp", names(chicago))
head(chicago[, -(i:j)])
```


## filter

```{r}
chic.f <- filter(chicago, pm25tmean2 > 30)
head(select(chic.f, 1:3, pm25tmean2), 10)
```

```{r}
chic.f %>%
  select(1:3) %>%
  head(.,10)
```

```{r}

chic.f %>%
  select(1:3,pm25tmean2) %>%
  head(.,10)

```


```{r}
chic.f2 <- chicago %>% 
  filter(pm25tmean2 > 30 & tmpd > 80)
```

```{r}
chic.f2 %>%
  select(1:3,pm25tmean2,tmpd) %>% 
  head(10)
```

## arrange 

Reordering rows of a data frame (while preserving corresponding order of other columns) is normally a pain to do in R

```{r}
library(dplyr)
```


```{r}
chicago <- readRDS("data/chicago.rds")
dim(chicago)
chicago <- arrange(chicago, date)
```

```{r}
head(select(chicago, date, pm25tmean2), 3)
```

```{r}
chicago %>%
  select(date, pm25tmean2) %>% 
  head(.,3)
```

```{r}
tail(select(chicago, date, pm25tmean2), 3)
```

```{r}
chicago %>%
  select(date, pm25tmean2) %>% 
  tail(.,3)
```

Columns can be arranged in descending order too. 

```{r}
chicago <- arrange(chicago, desc(date)) 
head(select(chicago, date, pm25tmean2), 3)
```

```{r}
chicago %>%
  arrange(desc(date)) %>%
  select(date, pm25tmean2) %>% 
  head()
```

```{r}
tail(select(chicago, date, pm25tmean2), 3)
```


## rename

Renaming a variable in a data frame in R is surprising hard to do!

```{r}
head(chicago[,1:5],3)
```

```{r}
chicago <- rename(chicago, dewpoint = dptp, pm25 = pm25tmean2)
head(chicago[,1:5],3)
```
```{r,error=TRUE}
chicago <- chicago %>%
  rename(dewptp = dewpoint)
```

```{r}
head(chicago[,1:5])
```

## mutate

```{r}
chicago <- chicago %>%
  mutate(pm25detrend = pm25 - mean(pm25, na.rm = TRUE))
```

```{r}
chicago %>% 
  select(pm25,pm25detrend) %>%
  head(.)
```


## group_by

Generating summary statistics by stratum 

```{r}
chicago <- chicago %>%
  mutate(tempcat = factor(1 * (tmpd >80), labels = c("cold","hot")))
```


```{r}
hotcold <- chicago %>% 
  group_by(tempcat)
```

```{r}
summarise(hotcold, pm25 = mean(pm25, na.rm = TRUE), o3 = max(o3tmean2), no2 = median(no2tmean2))
```

Generating summary statistics by stratum

```{r}
chicago <- mutate(chicago, year = as.POSIXlt(date)$year + 1900)
years <- group_by(chicago, year)
```

```{r}
years
```

```{r}
summarize(years, pm25 = mean(pm25, na.rm = TRUE),
o3 = max(o3tmean2, na.rm = TRUE),
no2 = median(no2tmean2, na.rm = TRUE))
```

## `%>%`

```{r}
chicago %>% 
  mutate(month = as.POSIXlt(date)$mon +1) %>%
  group_by(month) %>%
  summarise(pm25=mean(pm25, na.rm = TRUE),
            o3 = max(o3tmean2, na.rm = T),
            no2 = median(no2tmean2, na.rm = T))
```

## dplyr 


dplyr
Once you learn the dplyr “grammar” there are a few additional benefits
dplyr can work with other data frame “backends” data.table for large fast tables
SQL interface for relational databases via the DBI package

