---
title: "Merging Data"
output:
  pdf_document:
    toc: yes
    toc_depth: 6
    number_sections: yes
    latex_engine: xelatex
    df_print: kable
    highlight: tango
  html_notebook: default
  html_document:
    toc: yes
    toc_depth: '6'
    df_print: paged
  word_document:
    toc: yes
    toc_depth: '6'
---

# Merging Data

## Peer review data

```{r}
if(!file.exists("data")){dir.create("data")}
```


```{r}
reviews = read.csv("./data/reviews.csv"); 
solutions <- read.csv("./data/solutions.csv")
head(reviews,2)
```

```{r}
head(solutions,2)
```


## Merging data - merge()

* Merges data frames
* Important parameters: _x_,_y_,_by_,_by.x_,_by.y_,_all_

```{r}
names(reviews)
```

```{r}
names(solutions)
```

```{r}
mergedData <- merge(reviews, solutions, by.x = "solution_id", by.y = "id", all = TRUE)
```


```{r}
head(mergedData)
```


## Default - merge all common column names


```{r}
intersect(names(reviews),names(solutions))
```

```{r}
mergedData2 = merge(reviews,solutions,all=TRUE)
head(mergedData2)
```

## Using join in the plyr package

```{r}
library(dplyr)
```

```{r}
df1 <- data.frame(id=sample(1:10), x=rnorm(10))
df2 <- data.frame(id=sample(1:10), y=rnorm(10))
```


```{r}
head(df1)
```

```{r}
head(df2)
```


_Faster, but less full featured - defaults to left join, see help file for more_

```{r}
library(plyr)
```


```{r}
arrange(join(df1,df2),id)
```

## If you have multiple data frames

```{r}
df1 = data.frame(id=sample(1:10),x=rnorm(10))
df2 = data.frame(id=sample(1:10),y=rnorm(10))
df3 = data.frame(id=sample(1:10),z=rnorm(10))
dfList = list(df1,df2,df3)
join_all(dfList)
```

## More on merging data

* The quick R data merging page - [http://www.statmethods.net/management/merging.html](http://www.statmethods.net/management/merging.html)
* plyr information - [http://plyr.had.co.nz/](http://plyr.had.co.nz/)
* Types of joins - [http://en.wikipedia.org/wiki/Join_(SQL)](http://en.wikipedia.org/wiki/Join_(SQL))