---
title: "Subsetting and Sorting"
output:
  pdf_document:
    toc: yes
    toc_depth: 6
    number_sections: yes
    latex_engine: xelatex
    df_print: kable
    highlight: tango
  html_notebook: default
  html_document:
    toc: yes
    toc_depth: '6'
    df_print: paged
  word_document:
    toc: yes
    toc_depth: '6'
---

# Subsetting and Sorting

## Subsetting -quick review

```{r}
set.seed(13435)
X <- data.frame("var1"=sample(1:5),"var2"=sample(6:10),"var3"=sample(11:15))
X
```

```{r}
X <- X[sample(1:5),] # kind with random order?
X
```

```{r}
X$var2[c(1,3)] = NA
X
```

```{r}
X[,1] # get first col 
```

```{r}
X[,"var1"] # get col with name
```

```{r}
X[1:2,"var2"] # get 1st and snd row of col named "var2"
```

## Logicals ands and ors

```{r}
X[(X$var1 <= 3 & X$var3 > 11),] # get all rows that meets the condition
```

```{r}
X[(X$var1 <=3 | X$var3 > 15),] # logical or 
```

## Dealing with missing values

```{r}
X[which(X$var2 >8),]
```

## Sorting 

```{r}
sort(X$var1)
```

```{r}
sort(X$var1, decreasing = T)
```

```{r}
sort(X$var2, na.last=TRUE) # put NA values at the last 
```

## Ordering 

```{r}
X[order(X$var1),]
```

```{r}
X[order(X$var1,X$var3),]
```


The order() function in R returns a permutation of the vector that rearranges it into ascending or sorted order.

So, X[order(X$var1, X$var3),] is reordering the rows of the data frame X first by the values of var1 in ascending order. For rows where var1 is equal, it will then order by var3.

The result will be a data frame where the rows are sorted first by var1, and then by var3. 

## Ordering with plyr 

```{r}
library(plyr)
```

The pieces are the "ply" (similar to how plywood is layers of wood veneer) and the "r" represents the R programming language.

The plyr package is a set of tools for manipulating data in R. It provides a number of functions that make it easier to manipulate and reshape data.


```{r}
arrange(X,var1)
```

```{r}
arrange(X,desc(var1))
```

## Adding rows and columns

```{r}
X$var4 <- rnorm(5)
X
```

```{r}
Y <- cbind(X,rnorm(5))
Y
```

## Notes and further resources

R programming in the Data Science Track
Andrew Jaffe's lecture notes 1http://www.biostat.jhsph.edu/~ajaffe/lec_winterR/Lecture%202.pdf

