---
title: "Graphs"
output:
pdf_document:
toc: yes
toc_depth: 6
number_sections: yes
latex_engine: xelatex
df_print: kable
highlight: tango
html_notebook: default
html_document:
toc: yes
toc_depth: '6'
df_print: paged
word_document:
toc: yes
toc_depth: '6'
---

# Graphs


```{r readpm25data}
pollution <- read.csv("data/avgpm25.csv", colClasses = c("numeric", "character", "factor", "numeric", "numeric"))
names(pollution)
```

```{r}
str(pollution) ; str(pollution)
```

```{r}
visdat::vis_dat(pollution)
```

```{r}
visdat::vis_miss(pollution)
```
```{r}
boxplot(pollution$pm25, col = "blue")
```
```{r}
hist(pollution$pm25, col = "green")
```
```{r}
hist(pollution$pm25, col = "green", breaks = 100)
rug(pollution$pm25)
```
```{r}
boxplot(pollution$pm25, col = "blue")
abline(h=12)
```
```{r}
cat("aa")
```
```{r}
barplot(table(pollution$region),col = "wheat", main = "Number of Countries in Each Region")
```
```{r}
table(pollution$region)
```
```{r}
boxplot(pm25 ~ region, data = pollution, col = "red")
```

par can be used to set or query graphical parameters. Parameters can be set by specifying them as arguments to par in tag = value form, or by passing them as a list of tagged values.

mfcol, mfrow
A vector of the form c(nr, nc). Subsequent figures will be drawn in an nr-by-nc array on the device by columns (mfcol), or rows (mfrow), respectively.
In a layout with exactly two rows and columns the base value of "cex" is reduced by a factor of 0.83: if there are three or more of either rows or columns, the reduction factor is 0.66.



```{r}
par(mfrow = c(2,1), mar = c(4,4,2,1)) # 2 row 1 col
hist(subset(pollution, region = "west")$pm25, col = "green")
hist(subset(pollution, region = "east")$pm25, col = "green")
```

