---
title: "R Notebook"
output: html_notebook
---

## Matrix data

```{r}
set.seed(123)
par(mar = rep(0, 2, 4))
dataMatrix <- matrix(rnorm(400), nrow = 40)
image(1:10, 1:40, t(dataMatrix)[, nrow(dataMatrix):1])
```


## cluster the data



```{r, error = TRUE}
par(mar = rep(0.2, 4))
heatmap(dataMatrix)
```


## What if we add a pattern ?
```{r}
set.seed(678910)
for (i in 1:40) {
  # flip a coin
  coinFlip <- rbinom(1, size = 1, prob = 0.5)
  # if coin is heads add a common pattern to that row if (coinFlip) {
  dataMatrix[i,] <- dataMatrix[i,] + rep(c(0, 3), each = 5)
}
```
```{r}
par(mar = rep(0.2, 4))
image(1:10, 1:40, t(dataMatrix)[, nrow(dataMatrix):1])
```
```{r}
par(mar = rep(0.2, 4))
heatmap(dataMatrix)
```
## Patterns in rows and columns

```{r}
hh <- hclust(dist(dataMatrix))
dataMatrixOrdered <- dataMatrix[hh$order, ]
par(mfrow = c(1, 3))
image(t(dataMatrixOrdered)[, nrow(dataMatrixOrdered):1])
plot(rowMeans(dataMatrixOrdered), 40:1, , xlab = "Row Mean", ylab = "Row", pch = 19)
plot(colMeans(dataMatrixOrdered), xlab = "Column", ylab = "Column Mean", pch = 19)
```
## Related problems

You have multivariate variables X1,...,Xn so X1 = (X11,...,X1m)
· Find a new set of multivariate variables that are uncorrelated and explain as much variance as
possible.
· If you put all the variables together in one matrix, find the best matrix created with fewer variables (lower rank) that explains the original data.
The first goal is statistical and the second goal is data compression.


a lot of X might be correlated to each other, so we want to create a set of variables that reduced but able to explain
as much as possible the original data.


## Related solutions - PCA/SVD

SVD
If X is a matrix with each variable in a column and each observation in a row then the SVD is a "matrix decomposition"
$X = UDV^T$
where the columns of U are orthogonal (left singular vectors), the columns of V are orthogonal (right
singular vectors) and D is a diagonal matrix (singular values).

PCA
The principal components are equal to the right singular values if you first scale (subtract the mean, divide by the standard deviation) the variables.

