---
title: "R Notebook"
output: html_notebook
---
```{r}
library(ggplot2)
str(mpg)
```
```{r}
qplot(displ, hwy, data = mpg)
```
```{r}
qplot(displ, hwy, data = mpg, color = drv)
```
```{r}
qplot(displ,hwy, data = mpg,geom = c("point", "smooth"))
```
```{r}
qplot(hwy, data = mpg, fill = drv)
```
```{r}
qplot(displ, hwy, data = mpg, facets = .~ drv)
```
```{r}
qplot(hwy, data = mpg, facets = drv ~ ., bindwidth = 0.2)
```


ggplot2 -- grammar of graphics

- dataframe
- aesthetic mappings how data are mappped to color size shape
- geoms geometric objects like points lines, shapes
- facets : for conditional plots
- stats: statistical transformations like binning, quantiles, smoothing
- scales: scales for x and y axis, color, size etc
- coordinate system: the system that describes the 2d space that data lives in


- plots are built up in layers
- plot the data
- overlay a summary
- metadata and annotations
```{r}
ggplot(data = mpg, aes(x=cty,y=hwy))
```
```{r}
ggplot(data = mpg, aes(x = cty, y = hwy)) + geom_point()
```
```{r}
testdat <- data.frame(x = 1:100, y=rnorm(100))
testdat[50,2] <- 100 # outliers
plot(testdat$x, testdat$y, type = "l", ylim = c(-3,3))
```
```{r}
g <- ggplot(testdat, aes(x,y))
```
```{r}
g + geom_point()
```
```{r}
g + geom_line()
```
```{r}
g + geom_point() + geom_line()
```
```{r}
g + geom_point() + geom_line() + ylim(-3,3)
```
```{r}
```
