---
title: "R Notebook"
output:
  pdf_document:
    toc: yes
    toc_depth: 6
    number_sections: yes
    latex_engine: xelatex
    df_print: kable
    highlight: tango
  html_notebook: default
  html_document:
    toc: yes
    toc_depth: '6'
    df_print: paged
  word_document:
    toc: yes
    toc_depth: '6'
---

# Programming Assignment 2 Lexical Scoping

second programming assignment will require you to write an R function is able to cache potentially time-consuming computations. For example, taking the mean of a numeric vector is typically a fast operation. However, for a very long vector, it may take too long to compute the mean, especially if it has to be computed repeatedly (e.g. in a loop). If the contents of a vector are not changing, it may make sense to cache the value of the mean so that when we need it again, it can be looked up in the cache rather than recomputed. In this Programming Assignment will take advantage of the scoping rules of the R language and how they can be manipulated to preserve state inside of an R object.


## Example: Caching the Mean of a Vector

In this example we introduce the <<- operator which can be used to assign a value to an object in an environment that is different from the current environment. Below are two functions that are used to create a special object that stores a numeric vector and cache's its mean. 

The first function, makeVector creates a special "vector", which is really a list containing a function to

set the value of the vector

get the value of the vector

set the value of the mean

get the value of the mean

```{r}
makeVector <- function(x = numeric()) {
  # an object that will hold the cached mean of the vector x. initially NULL
  m <- NULL
  
  set <- function(y) {
    x <<- y # update the value of x 
    m <<- NULL # invalidate the cache (set to NULL)
  }
  
  get <- function() x # returns the current vector x
  
  setmean <- function(mean) m <<- mean # updates the cached mean m 
  
  getmean <- function() m # returns the cached mean m.
  
  #  Finally, makeVector returns a list of the set, get, setmean, and getmean functions. 
  # This list represents a special kind of vector that 
  # has its own methods for getting and setting the vector and its mean.
  
  list(
    set = set,
    get = get,
    setmean = setmean,
    getmean = getmean
  )
  
}
```

This pattern is very similar to the concept of Object-Oriented Programming (OOP) in other languages. In OOP, an object is an instance of a class, and a class can have properties (data attributes) and methods (functions) associated with it.

In this case, the function makeVector is akin to a class constructor that creates an object (the list of functions). The object has properties (the vector x and the mean m) that are "private" due to R's lexical scoping rules. The object also has methods (the functions set, get, setmean, and getmean) that provide a controlled interface for interacting with its properties.

This approach in R isn't strictly OOP as in languages like Java or Python. R does support more formal OOP paradigms, including S3, S4, and reference classes, but this functional approach is a popular alternative that leverages R's lexical scoping rules to achieve encapsulation, one of the key features of OOP.

So while R doesn't follow the same OOP syntax as languages like Java or Python, this pattern achieves a similar result: an object that bundles together data and the functions that operate on that data, with some control over the visibility and accessibility of the object's internal state.

The following function calculates the mean of the special "vector" created with the above function. However, it first checks to see if the mean has already been calculated. If so, it gets the mean from the cache and skips the computation. Otherwise, it calculates the mean of the data and sets the value of the mean in the cache via the setmean function.

```{r}
cachemean <- function(x, ...) {
  m <- x$getmean()
  if (!is.null(m)) {
    message("getting cached data")
    return(m)
  }
  data <- x$get()
  m <- mean(data, ...)
  x$setmean(m)
  m
}
```

The cachemean function is designed to work with an object that has been created using the makeVector function, but it doesn't call makeVector directly. Instead, we would need to call makeVector to create a special vector object first, and then you would pass this object to the cachemean function. 

That is to say, the makeVector is a class, we need to initialize it and give instance of makeVector to cachemean. 

The makeVector function could be thought of as a constructor function, similar to how you might initialize a new object from a class in other languages. It creates an instance of this special "vector" object, which has certain properties (the actual numeric vector x and its mean m) and methods (set, get, setmean, and getmean).

Then, the cachemean function is designed to work with these special "vector" objects. It expects an instance of a makeVector object as its input, and it uses the methods provided by this object to get the vector data, get the cached mean, and set the cached mean.

This kind of programming style is known as functional programming. It's different from traditional object-oriented programming, but there are some similarities. In this case, the behavior and state of the "vector" objects are defined by the makeVector function and the functions it encapsulates, and the cachemean function is designed to interact with these objects in a specific way.


```{r}
# create a special vector object
my_vector <- makeVector(1:10)

# calculate and cache the mean of the vector
mean_value <- cachemean(my_vector)

mean_value
```

```{r}
# print the calculated mean
print(mean_value)
```
 my_vector is stored in the R environment in which you're working. In this case, it's stored in your current workspace. It's assigned the return value of the makeVector function, which is a list of functions (set, get, setmean, and getmean).

The my_vector object, like all R objects in current session, is saved in your computer's memory (RAM), not on the disk. This is because memory storage allows for much quicker access and modification of the object.

## Assignment: Caching the Inverse of a Matrix

Matrix inversion is usually a costly computation and there may be some benefit to caching the inverse of a matrix rather than compute it repeatedly (there are also alternatives to matrix inversion that we will not discuss here). Your assignment is to write a pair of functions that cache the inverse of a matrix.

Write the following functions:

*makeCacheMatrix*: This function creates a special "matrix" object that can cache its inverse.

*cacheSolve*: This function computes the inverse of the special "matrix" returned by makeCacheMatrix above. If the inverse has already been calculated (and the matrix has not changed), then the cachesolve should retrieve the inverse from the cache.

Computing the inverse of a square matrix can be done with the solve function in R. For example, if X is a square invertible matrix, then solve(X) returns its inverse.

For this assignment, assume that the matrix supplied is always invertible.

```{r}
# This function creates a special "matrix" object that can cache its inverse.
# this is a blueprint or anomy of Object in OOP that has it's param and functions
makeCacheMatrix <- function(x = matrix()) {
  inv <- NULL # var to hold the inverse value of the matrix
  
  set <- function(new_val) {
    x <- new_val # update the new value
    inv <- NULL # invalidate the old one
  }
  
  get <- function() {
    x # get the value of x
  }
  
  setInverse <- function(val) {
    inv <- val # set the inverse value
  }
  
  getInverse <- function() {
    inv # get the value of inv
  }
  
  # return object methods (OOP static methods to call)
  list(
    set = set,
    get = get,
    setInverse = setInverse,
    getInverse = getInverse
  )
}
```

```{r}
# This function is used to work with matrix that generated by makeCacheMatrix.
# (Analogy of OOP Instantiation of the object)
# If the inveser already been calcualted, and the matrix is the same,
# then we just return the cached value instead of recalculating it.

cacheSolve <- function(x, ...) {
  inv <- x$getInverse() # try to retrieve the cached inverse
  
  if(!is.null(inv)) {
    # we are able to retrieve inverse from cache
    return(inv)
  }
  
  # cache dose not have the inverse
  data <- x$get() # get the actual matrix data that stored in the x
  inv <- solve(data, ...) # If you pass a square matrix A to the solve() function like solve(A)
  x$setInverse(inv) # then we want to cache the calculation result
  return(inv)
}
```

```{r}
B <- matrix(c(1,2,3,4),2,2)
B1 <- makeCacheMatrix(B)
cacheSolve(B1)
```

## Testing 

```{r}
A = matrix(c(4, 7, 2, 6), nrow = 2)
A
```

```{r}
solve(A)
```

```{r}
cacheSolve(
  makeCacheMatrix(A)
)
```

