---
title: "R Notebook"
output: html_notebook
---

# Control Structures 



## Control Structures - Introduction

Control structures in R allow you to control the flow of execution of the program, depending on runtime conditions. Common structures are
· if, else: testing a condition
· for: execute a loop a fixed number of times
· while: execute a loop while a condition is true · repeat: execute an infinite loop
· break: break the execution of a loop
· next: skip an interation of a loop
· return: exit a function
Most control structures are not used in interactive sessions, but rather when writing functions or longer expresisons.

## Control Structures - If-else

```
if(<condition>) {
## do something
}else{
## do something else
} if(<condition1>) {
## do something
} else if(<condition2>) {
## do something different
}else{
## do something different
}
```
### if

This is a valid if/else structure.
```{r}
x <-0
```


```{r}
if(x>3) {
  y <- 10
} else{
  y <- 0
}
```

So is this one.

```{r}
y <- if (x > 3) {
  10
} else{
  0
}
```

Of course, the else clause is not necessary.

```
if(<condition1>) { 

}

if(<condition2>) {

}
```


