---
title: "R Notebook"
output: html_notebook
---

# Control Structures 



## Control Structures - Introduction

Control structures in R allow you to control the flow of execution of the program, depending on runtime conditions. Common structures are
· if, else: testing a condition
· for: execute a loop a fixed number of times
· while: execute a loop while a condition is true · repeat: execute an infinite loop
· break: break the execution of a loop
· next: skip an interation of a loop
· return: exit a function
Most control structures are not used in interactive sessions, but rather when writing functions or longer expresisons.

## Control Structures - If-else

```
if(<condition>) {
## do something
}else{
## do something else
} if(<condition1>) {
## do something
} else if(<condition2>) {
## do something different
}else{
## do something different
}
```
### if

This is a valid if/else structure.
```{r}
x <-0
```


```{r}
if(x>3) {
  y <- 10
} else{
  y <- 0
}
```

So is this one.

```{r}
y <- if (x > 3) {
  10
} else{
  0
}
```

Of course, the else clause is not necessary.

```
if(<condition1>) { 

}

if(<condition2>) {

}
```

## Control Structures - For loops

for loops take an interator variable and assign it successive values from a sequence or vector. For loops are most commonly used for iterating over the elements of an object (list, vector, etc.)

```{r}
for(i in 1:10) {
  print(i)
}
```

This loop takes the i variable and in each iteration of the loop gives it values 1, 2, 3, ..., 10, and then exits.

These three loops have the same behavior.

```{r}
x <- c("a", "b", "c", "d")

for (i in 1:4) {
  print(x[i])
}
```


```{r}
for(i in seq_along(x)) {
  print(x[i])
}
```

```{r}
for(letter in x) {
  print(letter)
}
```

```{r}
for(i in 1:4) print(x[i])
```

### Nested for loops

for loops can be nested.

```{r}
x <- matrix(1:6, 2, 3)
for (i in seq_len(nrow(x))) {
  for (j in seq_len(ncol(x))) {
    print(x[i, j])
  }
}
```

Be careful with nesting though. Nesting beyond 2–3 levels is often very difficult to read/understand.