---
title: "Programming Assignment 1"
output: html_notebook
---

# Programming Assignment 1

```{r}
print(R.version.string)
```


## Part 1

Write a function named 'pollutantmean' that calculates the mean of a pollutant (sulfate or nitrate) across a specified list of monitors. The function 'pollutantmean' takes three arguments: 'directory', 'pollutant', and 'id'. Given a vector monitor ID numbers, 'pollutantmean' reads that monitors' particulate matter data from the directory specified in the 'directory' argument and returns the mean of the pollutant across all of the monitors, *ignoring any missing values* coded as NA. A prototype of the function is as follows


```{r}
files <- list.files("specdata")
files
```

```{r}
files[1]
```
```{r}
files[1:10]
```


```{r}
pollutantmean <- function(directory, pollutant, id = 1:332) {
  df <- data.frame()
  files <- list.files(directory, full.names = TRUE)
  for (i in id) {
    tmp.file <- files[i]
    tmp.df <- read.csv(tmp.file)
    df <- rbind(df, tmp.df)
  }
  return(mean(df[[pollutant]], na.rm = TRUE))
}
```

```{r}
pollutantmean("specdata", "sulfate", 1)
```

```{r}
pollutantmean("specdata", "sulfate", 1:10)
```

```{r}
pollutantmean("specdata", "nitrate", 70:72)
```

```{r}
pollutantmean("specdata", "nitrate", 23)
```

## Part 2

```{r}
df <- data.frame(id=integer(),nobs=integer())
```


```{r}
files <- list.files("specdata", full.names = TRUE)
```

```{r}
i <- 1
```

```{r}
tmp.file <- files[i]
tmp.df <- read.csv(tmp.file)
head(tmp.df)
```
```{r}
nobs <- sum(complete.cases(tmp.df))
nobs
```

```{r}
df <- rbind(df, data.frame(id = i, nobs = nobs))
df
```




```{r}
complete <- function(directory, id = 1:332) {
  df <- data.frame(id = integer(), nobs = integer())
  files <- list.files(directory, full.names = TRUE)
  for (i in id) {
    tmp.file <- files[i]
    tmp.df <- read.csv(tmp.file)
    nobs <- sum(complete.cases(tmp.df))
    df <- rbind(df, data.frame(id = i, nobs = nobs))
  }
  return(df)
}
```

```{r}
complete("specdata", 1)
```
```{r}
complete("specdata", c(2, 4, 8, 10, 12))
```

```{r}
complete("specdata", 30:25)
```

```{r}
complete("specdata", 3)
```

## Part 3

```{r}
files <- list.files("specdata", full.names = TRUE)
files[1]
```
```{r}
i=1
threshold = 10
```

```{r}
tmp.file <- files[i]
tmp.file
```
```{r}
tmp.df <- read.csv(tmp.file)
```


```{r}
corr <- function(directory, threshold = 0) {
  corrs <- c()
  files <- list.files(directory, full.names = TRUE)
  for (i in id) {
    tmp.file <- files[i]
    tmp.df <- read.csv(tmp.file)
    nobs <- sum(complete.cases(tmp.df))
    if (nobs > threshold) {
      corrs <- c(corrs, cor(data$sulfate, data$nitrate, use = "complete.obs"))
    }
  }
  return(corrs)
}
```



```{r}
corr("specdata", 10)
```
```{r}
corr <- function(directory, threshold = 0) {
    # Initialize an empty vector to hold the correlation coefficients
    corrs <- c()
    
    # Get the list of file names in the directory
    file_names <- list.files(directory)
    
    # Loop through each file name
    for(file_name in file_names) {
        # Construct the file path
        file_path <- paste(directory, "/", file_name, sep = "")
        
        # Read the data from the file
        data <- read.csv(file_path)
        
        # Check if the number of complete cases is greater than the threshold
        if(sum(complete.cases(data)) > threshold) {
            # If so, calculate the correlation coefficient and add it to the corrs vector
            corrs <- c(corrs, cor(data$sulfate, data$nitrate, use = "complete.obs"))
        }
    }
    
    # Return the corrs vector
    return(corrs)
}
```

```{r}
cr <- corr("specdata", 150)
```

```{r}
head(cr)
```
```{r}
summary(cr)
```
```{r}
cr <- corr("specdata", 400)
head(cr)
```



```{r}
cr <- corr("specdata", 5000)
summary(cr)
```

```{r}
length(cr)
```

```{r}
pollutantmean("specdata", "sulfate", 1:10)
```

```{r}
pollutantmean("specdata", "nitrate", 70:72)
```

```{r}
pollutantmean("specdata", "sulfate", 34)
```

```{r}
pollutantmean("specdata", "nitrate")
```

```{r}
cc <- complete("specdata", c(6, 10, 20, 34, 100, 200, 310))
print(cc$nobs)
```

```{r}
cc <- complete("specdata", 54)
print(cc$nobs)
```

```{r}
RNGversion("3.5.1")  
set.seed(42)
cc <- complete("specdata", 332:1)
use <- sample(332, 10)
print(cc[use, "nobs"])
```

```{r}
cr <- corr("specdata")                
cr <- sort(cr)   
RNGversion("3.5.1")
set.seed(868)                
out <- round(cr[sample(length(cr), 5)], 4)
print(out)
```

```{r}
cr <- corr("specdata", 129)                
cr <- sort(cr)                
n <- length(cr)    
RNGversion("3.5.1")
set.seed(197)                
out <- c(n, round(cr[sample(n, 5)], 4))
print(out)
```

```{r}
cr <- corr("specdata", 2000)                
n <- length(cr)                
cr <- corr("specdata", 1000)                
cr <- sort(cr)
print(c(n, round(cr, 4)))
```

