---
title: "Programming Assignment 3 INSTRUCTIONS - Hospital Quality"
output:
  pdf_document:
    toc: yes
    toc_depth: 6
    number_sections: yes
    latex_engine: xelatex
    df_print: kable
    highlight: tango
  html_notebook: default
  html_document:
    toc: yes
    toc_depth: '6'
    df_print: paged
  word_document:
    toc: yes
    toc_depth: '6'
---

# Programming Assignment 3 INSTRUCTIONS - Hospital Quality

## data

### 11. Hospital_Data.csv

The Hospital_Data.csv table contains thirteen (13) fields. This table provides general Hospital information in response to a Hospital Compare search.
1. Provider Number: varchar (6) Lists the hospitals by their provider identification number. 2. Hospital Name: varchar (50) Lists the name of the hospital.
3. Address 1: varchar (50) Lists the first line of the street address of the hospital.
4. Address 2: varchar (50) Lists the second line of the street address of the hospital.
5. Address 3: varchar (50) Lists the third line of the street address of the hospital. 6. City: varchar (28) Lists the city in which the hospital is located.
7. State: varchar (2) Lists the 2 letter State code in which the hospital is located. 8. ZIP Code: char (5) Lists the 5 digit numeric ZIP for the hospital.
9. County Name: char (15) Lists the county in which the hospital is located.
10. Phone Number: char (10) Lists the 10-digit numeric telephone number, including area code, for the Hospital.
11. Hospital Type: char (25) Lists the type of hospital. The values are: a. Acute Care Hospital
b. Acute Care – VA Medical Center c. Critical Access Hospital
d. Children’s Hospital
12. Hospital Owner: varchar (44) Lists the type of ownership the Hospital falls under. The values are:
a. Government – Federal
b. Government – Hospital District or Authority c. Government – Local
d. Government – State
e. Proprietary
f. Voluntary non-profit – Church
g. Voluntary non-profit – Other
h. Voluntary non-profit – Private
i. Not Available
13. Emergency Services: char (3) Returns “Yes” or “No” to specify whether or not the hospital provides emergency services.

### 19. Outcome of Care Measures.csv 

The Outcome of Care Measures.csv table contains forty seven (47) fields. This table provides each hospital’s risk-adjusted 30-Day Death (mortality) and 30-Day Readmission category and rate.
1. Provider Number: varchar (6) Lists the hospitals by their provider identification number.
2. Hospital Name: varchar (50) Lists the name of the hospital.
3. Address 1: varchar (50) Lists the first line of the street address of the hospital.
4. Address 2: varchar (50) Lists the second line of the street address of the hospital.
5. Address 3: varchar (50) Lists the third line of the street address of the hospital.
6. City: varchar (28) Lists the city in which the hospital is located.
7. State: varchar (2) Lists the 2 letter State code in which the hospital is located.
8. ZIP Code: char (5) Lists the 5 digit numeric ZIP for the hospital.
9. County Name: char (15) Lists the county in which the hospital is located.
10. Phone Number: char (10) Lists the 10-digit numeric telephone number, including area code, for the Hospital.
11. Hospital 30-Day Death (Mortality) Rates from Heart Attack: Lists the risk adjusted rate (percentage) for each hospital.
12. Comparison to U.S. Rate - Hospital 30-Day Death (Mortality) Rates from Heart Attack: varchar (50) Lists the mortality and readmission category in which the hospital falls. The values are:
• Better than U.S. National Average
• No Different than U.S. National Average • Worse than U.S. National Average
• Number of Cases too Small*
13. Lower Mortality Estimate - Hospital 30-Day Death (Mortality) Rates from Heart Attack: Lists the lower bound (Interval Estimate) for each hospital’s risk-adjusted rate.
14. Upper Mortality Estimate - Hospital 30-Day Death (Mortality) Rates from Heart Attack: Lists the upper bound (Interval Estimate) for each hospital’s risk-adjusted rate.
15. Number of Patients - Hospital 30-Day Death (Mortality) Rates from Heart Attack: varchar (5) Lists the number of Medicare patients treated for Heart Attack by the Hospital.
16. Footnote - Hospital 30-Day Death (Mortality) Rates from Heart Attack: Lists the footnote value when appropriate, as related to the Heart Attack Outcome of Care at the hospital.
17. Hospital 30-Day Death (Mortality) Rates from Heart Failure: Lists the risk adjusted rate (percentage) for each hospital.
18. Comparison to U.S. Rate - Hospital 30-Day Death (Mortality) Rates from Heart Failure: varchar (50) Lists the mortality and readmission category in which the hospital falls. The values are:
a. Better than U.S. National Average
b. No Different than U.S. National Average c. Worse than U.S. National Average
d. Number of Cases too Small*
19. Lower Mortality Estimate - Hospital 30-Day Death (Mortality) Rates from Heart Failure: Lists the lower bound (Interval Estimate) for each hospital’s risk-adjusted rate.
20. Upper Mortality Estimate - Hospital 30-Day Death (Mortality) Rates from Heart Failure: Lists the upper bound (Interval Estimate) for each hospital’s risk-adjusted rate.
21. Number of Patients - Hospital 30-Day Death (Mortality) Rates from Heart Failure: varchar (5) Lists the number of Medicare patients treated for Heart Failure by the Hospital.
22. Footnote - Hospital 30-Day Death (Mortality) Rates from Heart Failure: Lists the footnote value when appropriate, as related to the Heart Failure Outcome of Care at the hospital.
23. Hospital 30-Day Death (Mortality) Rates from Pneumonia: Lists the risk adjusted rate (percentage) for each hospital.
24. Comparison to U.S. Rate - Hospital 30-Day Death (Mortality) Rates from Pneumonia: varchar (50) Lists the mortality and readmission category in which the hospital falls. The values are:
• Better than U.S. National Average
• No Different than U.S. National Average • Worse than U.S. National Average
• Number of Cases too Small*
25. Lower Mortality Estimate - Hospital 30-Day Death (Mortality) Rates from Pneumonia: Lists the lower bound (Interval Estimate) for each hospital’s risk-adjusted rate.
26. Upper Mortality Estimate - Hospital 30-Day Death (Mortality) Rates from Pneumonia: Lists the upper bound (Interval Estimate) for each hospital’s risk-adjusted rate.
27. Number of Patients - Hospital 30-Day Death (Mortality) Rates from Pneumonia: varchar (5) Lists the number of Medicare patients treated for Pneumonia by the Hospital.
28. Footnote - Hospital 30-Day Death (Mortality) Rates from Pneumonia: Lists the footnote value when appropriate, as related to the Pneumonia Outcome of Care at the hospital.
29. Hospital 30-Day Readmission Rates from Heart Attack: Lists the risk adjusted rate (percentage) for each hospital.
30. Comparison to U.S. Rate - Hospital 30-Day Readmission Rates from Heart Attack: varchar (50) Lists the mortality and readmission category in which the hospital falls. The values are:
• Better than U.S. National Average
• No Different than U.S. National Average • Worse than U.S. National Average
• Number of Cases too Small*
32.Lower Readmission Estimate - Hospital 30-Day Readmission Rates from Heart Attack: Lists the lower bound (Interval Estimate) for each hospital’s risk-adjusted rate.
33.Upper Readmission Estimate - Hospital 30-Day Readmission Rates from Heart Attack: Lists the upper bound (Interval Estimate) for each hospital’s risk-adjusted rate.
34. Number of Patients - Hospital 30-Day Readmission Rates from Heart Attack: varchar (5) Lists the number of Medicare patients treated for Heart Attack.
35. Footnote - Hospital 30-Day Readmission Rates from Heart Attack: Lists the footnote value when appropriate, as related to the Heart Attack Outcome of Care at the hospital.
36. Hospital 30-Day Readmission Rates from Heart Failure: Lists the risk adjusted rate (percentage) for each hospital.
37. Comparison to U.S. Rate - Hospital 30-Day Readmission Rates from Heart Failure: varchar (50) Lists the mortality and readmission category in which the hospital falls. The values are:
• Better than U.S. National Average
• No Different than U.S. National Average • Worse than U.S. National Average
• Number of Cases too Small*
38.Lower Readmission Estimate - Hospital 30-Day Readmission Rates from Heart Failure: Lists the lower bound (Interval Estimate) for each hospital’s risk-adjusted rate.
39.Upper Readmission Estimate - Hospital 30-Day Readmission Rates from Heart Failure: Lists the upper bound (Interval Estimate) for each hospital’s risk-adjusted rate.
40. Number of Patients - Hospital 30-Day Readmission Rates from Heart Failure: varchar (5) Lists the number of Medicare patients treated for Heart Failure.
41. Footnote - Hospital 30-Day Readmission Rates from Heart Failure: Lists the footnote value when appropriate, as related to the Heart Failure Outcome of Care at the hospital.
42. Hospital 30-Day Readmission Rates from Pneumonia: Lists the risk adjusted rate (percentage) for each hospital.
43.Comparison to U.S. Rate - Hospital 30-Day Readmission Rates from Pneumonia: varchar (50) Lists the mortality and readmission category in which the hospital falls. The values are:
• Better than U.S. National Average
• No Different than U.S. National Average • Worse than U.S. National Average
• Number of Cases too Small*

44. Lower Readmission Estimate - Hospital 30-Day Readmission Rates from Pneumonia: Lists the lower bound (Interval Estimate) for each hospital’s risk-adjusted rate.
45. Upper Readmission Estimate - Hospital 30-Day Readmission Rates from Pneumonia: Lists the upper bound (Interval Estimate) for each hospital’s risk-adjusted rate.
46. Number of Patients - Hospital 30-Day Readmission Rates from Pneumonia: varchar (5) Lists the number of Medicare patients treated for Pneumonia.
47. Footnote - Hospital 30-Day Readmission Rates from Pneumonia: Lists the footnote value when appropriate, as related to the Pneumonia Outcome of Care at the hospital.

## 1 Plot the 30-day mortality rates for heart attack

Read the outcome data into R via the read.csv function and look at the first few rows.

```{r}
outcome <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
head(outcome)
```
```{r}
ncol(outcome)
```

```{r}
nrow(outcome)
```

```{r}
names(outcome)
```
To make a simple histogram of the 30-day death rates from heart attack (column 11 in the outcome dataset), run

```{r}
outcome[, 11] <- as.numeric(outcome[, 11])
```
You may get a warning about NAs being introduced; that is okay 

```{r}
hist(outcome[, 11])
```
Because we originally read the data in as character (by specifying colClasses = "character" we need to coerce the column to be numeric. You may get a warning about NAs being introduced but that is okay.

## 2 Finding the best hospital in a state

Write a function called best that take two arguments: the 2-character abbreviated name of a state and an outcome name. The function reads the outcome-of-care-measures.csv file and returns a character vector with the name of the hospital that has the best (i.e. lowest) 30-day mortality for the specified outcome in that state. The hospital name is the name provided in the Hospital.Name variable. The outcomes can be one of “heart attack”, “heart failure”, or “pneumonia”. Hospitals that do not have data on a particular outcome should be excluded from the set of hospitals when deciding the rankings.


Handling ties. If there is a tie for the best hospital for a given outcome, then the hospital names should be sorted in alphabetical order and the first hospital in that set should be chosen (i.e. if hospitals “b”, “c”, and “f” are tied for best, then hospital “b” should be returned).

The function should use the following template.

```
best <- function(state, outcome) {
        ## Read outcome data
}

```

The function should check the validity of its arguments. If an invalid state value is passed to best, the function should throw an error via the stop function with the exact message “invalid state”. If an invalid outcome value is passed to best, the function should throw an error via the stop function with the exact message “invalid outcome”.

```{r}
# The function reads the outcome-of-care-measures.csv file and
# returns a character vector with the name of the hospital
# that has the best (i.e. lowest) 30-day mortality
# for the specified outcome in that state.

best <- function(state, outcome) {
  ## Read outcome data
  outcome_data <-
    read.csv("outcome-of-care-measures.csv", colClasses = "character")
  
  ## Check that state and outcome are valid
  valid_states <- unique(outcome_data$State)
  valid_outcomes <-
    c(
      "heart attack" = "Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack",
      "heart failure" = "Hospital.30.Day.Death..Mortality..Rates.from.Heart.Failure",
      "pneumonia" = "Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia"
    )
  
  # validate state
  if (!(state %in% valid_states)) {
    stop("invalid state")
  }
  # validate outcome
  if (!(outcome %in% names(valid_outcomes))) {
    stop("invalid outcome")
  }
  
  ## Select data for the specific state and outcome
  state_data <- subset(outcome_data, State == state)
  state_data <-
    state_data[!is.na(state_data[[valid_outcomes[outcome]]]),]
  state_data[[valid_outcomes[outcome]]] <-
    as.numeric(state_data[[valid_outcomes[outcome]]])
  
  ## Return hospital name in that state with lowest 30-day death
  best_hospital <-
    state_data[which.min(state_data[[valid_outcomes[outcome]]]), "Hospital.Name"]
  return(best_hospital)
}
```


```{r}
best("TX", "heart attack")
```

```{r}
best("TX", "heart failure")
```

```{r}
best("MD", "heart attack")
```

```{r}
best("MD", "pneumonia")
```

```{r}
best("BB", "heart attack")
```

```{r}
best("NY", "hert attack")
```

## 3 Ranking hospitals by outcome in a state

Write a function called rankhospital that takes three arguments: the 2-character abbreviated name of a state (state), an outcome (outcome), and the ranking of a hospital in that state for that outcome (num). The function reads the outcome-of-care-measures.csv file and returns a character vector with the name of the hospital that has the ranking specified by the num argument. For example, the call
rankhospital("MD", "heart failure", 5)
would return a character vector containing the name of the hospital with the 5th lowest 30-day death rate for heart failure. The num argument can take values “best”, “worst”, or an integer indicating the ranking (smaller numbers are better). If the number given by num is larger than the number of hospitals in that state, then the function should return NA. Hospitals that do not have data on a particular outcome should be excluded from the set of hospitals when deciding the rankings.

Handling ties. It may occur that multiple hospitals have the same 30-day mortality rate for a given cause of death. In those cases ties should be broken by using the hospital name. For example, in Texas (“TX”), the hospitals with lowest 30-day mortality rate for heart failure are shown here.

```
head(texas)

                        Hospital.Name Rate Rank
3935       FORT DUNCAN MEDICAL CENTER  8.1    1
4085  TOMBALL REGIONAL MEDICAL CENTER  8.5    2
4103 CYPRESS FAIRBANKS MEDICAL CENTER  8.7    3
3954           DETAR HOSPITAL NAVARRO  8.7    4
4010           METHODIST HOSPITAL,THE  8.8    5
3962  MISSION REGIONAL MEDICAL CENTER  8.8    6

```
Note that Cypress Fairbanks Medical Center and Detar Hospital Navarro both have the same 30-day rate (8.7). However, because Cypress comes before Detar alphabetically, Cypress is ranked number 3 in this scheme and Detar is ranked number 4. One can use the order function to sort multiple vectors in this manner (i.e. where one vector is used to break ties in another vector).
The function should use the following template.

```
rankhospital <- function(state, outcome, num = "best") {
    ## Read outcome data
    ## Check that state and outcome are valid
    ## Return hospital name in that state with the given rank
    ## 30-day death rate
}

```
The function should check the validity of its arguments. If an invalid state value is passed to rankhospital, the function should throw an error via the stop function with the exact message “invalid state”. If an invalid outcome value is passed to rankhospital, the function should throw an error via the stop function with the exact message “invalid outcome”.

```{r}
# The function reads the outcome-of-care-measures.csv file and 
# returns a character vector with the name of the hospital 
# that has the ranking specified by the num argument.

rankhospital <- function(state, outcome, num = "best") {
  ## Read outcome data
  outcome_data <-
    read.csv("outcome-of-care-measures.csv", colClasses = "character")
  
  ## Check that state and outcome are valid
  valid_states <- unique(outcome_data$State)
  valid_outcomes <-
    c(
      "heart attack" = "Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack",
      "heart failure" = "Hospital.30.Day.Death..Mortality..Rates.from.Heart.Failure",
      "pneumonia" = "Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia"
    )
  
  if (!(state %in% valid_states)) {
    stop("invalid state")
  }
  
  if (!(outcome %in% names(valid_outcomes))) {
    stop("invalid outcome")
  }
  
  ## Select data for the specific state and outcome
  state_data <- subset(outcome_data, State == state)
  state_data <-
    state_data[!is.na(state_data[[valid_outcomes[outcome]]]),]
  state_data[[valid_outcomes[outcome]]] <-
    as.numeric(state_data[[valid_outcomes[outcome]]])
  
  ## Order the data
  state_data <-
    state_data[order(state_data[[valid_outcomes[outcome]]], state_data$Hospital.Name),]
  
  ## Handle num argument
  if (num == "best") {
    rank = 1
  } else if (num == "worst") {
    rank = nrow(state_data)
  } else if (is.numeric(num)) {
    if (num > nrow(state_data)) {
      return(NA)
    } else {
      rank = num
    }
  } else {
    stop("invalid num argument")
  }
  
  ## Return hospital name in that state with the given rank 30-day death rate
  return(state_data$Hospital.Name[rank])
}

```

```{r}
rankhospital("TX", "heart failure", 4)
```

```{r}
rankhospital("MD", "heart attack", "worst")
```

```{r}
rankhospital("MN", "heart attack", 5000)
```

## 4 Ranking hospitals in all states

Write a function called `rankall` that takes two arguments: an outcome name (outcome) and a hospital rank- ing (num). The function reads the outcome-of-care-measures.csv file and returns a 2-column data frame containing the hospital in each state that has the ranking specified in num. For example the function call rankall("heart attack", "best") would return a data frame containing the names of the hospitals that are the best in their respective states for 30-day heart attack death rates. The function should return a value for every state (some may be NA). The first column in the data frame is named hospital, which contains the hospital name, and the second column is named state, which contains the 2-character abbreviation for the state name. Hospitals that do not have data on a particular outcome should be excluded from the set of hospitals when deciding the rankings.
Handling ties. The rankall function should handle ties in the 30-day mortality rates in the same way that the rankhospital function handles ties.
The function should use the following template.

```
rankall <- function(outcome, num = "best") {
## Read outcome data
## Check that state and outcome are valid
## For each state, find the hospital of the given rank
## Return a data frame with the hospital names and the
## (abbreviated) state name
}
```
NOTE: For the purpose of this part of the assignment (and for efficiency), your function should NOT call the rankhospital function from the previous section.
The function should check the validity of its arguments. If an invalid outcome value is passed to rankall, the function should throw an error via the stop function with the exact message “invalid outcome”. The num variable can take values “best”, “worst”, or an integer indicating the ranking (smaller numbers are better). If the number given by num is larger than the number of hospitals in that state, then the function should return NA.


```{r}
# The function reads the outcome-of-care-measures.csv file and 
# returns a 2-column data frame containing the hospital in each state 
# that has the ranking specified in num.

rankall <- function(outcome, num = "best") {
  ## Read outcome data
  outcome_data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
  
  ## Check that outcome is valid
  valid_outcomes <- c("heart attack" = "Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack",
                      "heart failure" = "Hospital.30.Day.Death..Mortality..Rates.from.Heart.Failure",
                      "pneumonia" = "Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia")
  
  if (!(outcome %in% names(valid_outcomes))) {
    stop("invalid outcome")
  }
  
  ## Get list of states
  states <- unique(outcome_data$State)
  
  ## Initialize an empty data frame to hold results
  results <- data.frame(hospital = character(length(states)), state = states, stringsAsFactors = FALSE)
  
  ## For each state, find the hospital of the given rank
  for (i in seq_along(states)) {
    state_data <- subset(outcome_data, State == states[i])
    state_data <- state_data[!is.na(state_data[[valid_outcomes[outcome]]]), ]
    state_data[[valid_outcomes[outcome]]] <- as.numeric(state_data[[valid_outcomes[outcome]]])
    
    ## Order the data
    state_data <- state_data[order(state_data[[valid_outcomes[outcome]]], state_data$Hospital.Name), ]
    
    ## Handle num argument
    if (num == "best") {
      rank = 1
    } else if (num == "worst") {
      rank = nrow(state_data)
    } else if (is.numeric(num)) {
      if (num > nrow(state_data)) {
        rank = NA
      } else {
        rank = num
      }
    } else {
      stop("invalid num argument")
    }
    
    ## Add hospital name of the given rank to the results data frame
    results$hospital[i] <- state_data$Hospital.Name[rank]
  }
  
  ## Drop the rows with NA in the hospital column
  results <- na.omit(results)
  
  ## Sort the results by state
  results <- results[order(results$state), ]
  
  ## Return a data frame with the hospital names and the (abbreviated) state name
  return(results)
}

```

```{r}
head(rankall("heart attack", 20), 10)
```

```{r}
tail(rankall("pneumonia", "worst"), 3)
```

```{r}
tail(rankall("heart failure"), 10)
```

```{r}
best("SC", "heart attack")
```

```{r}
best("NY", "pneumonia")
```

```{r}
best("AK", "pneumonia")
```

```{r}
rankhospital("NC", "heart attack", "worst")
```

```{r}
library(data.table)

rankHospital <- function(state, outcome, num="best") {
    # Read outcome data
    dt <- data.table::fread("data/outcome-of-care-measures.csv")
    
    # change outcome to lowercase
    outcome <- tolower(outcome)
    
    # change variable name to prevent confusion
    chosen_state <- state

    # Check state and outcome are valid, if not return warning message
    if (!chosen_state %in% unique(dt[["State"]])) {
        stop("Invalid state")
    }
    if (!outcome %in% c("heart attack", "heart failure", "pneumonia")) {
        stop("Invalid outcome")
    }
    
    dt <- dt %>% 
        rename_with(~ tolower(gsub("^Hospital 30-Day Death \\(Mortality\\) Rates from ", "", .x))) %>%
        filter(state == chosen_state) %>%
        mutate(rate = suppressWarnings(as.numeric(get(outcome)))) %>%
        clean_names() %>%
        select(hospital_name, state, rate) %>%
        filter(complete.cases(.)) %>%
        arrange(rate, hospital_name) %>%
        mutate(rank = row_number())  

    if (num == "best") {
        unlist(head(dt[[1]], 1))
    }
    
    else if (num == "worst") {
        unlist(tail(dt[[1]], 1))
    }
    
    else {
        dt %>% 
            slice(num) %>%
            select(hospital_name) %>%
            unlist()
    }
}
```

```{r}
# create a rankhospital function
rankhospital <- function(state, outcome, rank = "best"){
  ## Read outcome data
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
  fd   <- as.data.frame(cbind(data[, 2],  # hospital
                              data[, 7],  # state
                              data[, 11],  # heart attack
                              data[, 17],  # heart failure
                              data[, 23]), # pneumonia
                        stringsAsFactors = FALSE)
  colnames(fd) <- c("hospital", "state", "heart attack", "heart failure", "pneumonia")
  
  ## Check that state and outcome are valid
  if (!state %in% fd[, "state"]) {
    stop('invalid state')
  } else if (!outcome %in% c("heart attack", "heart failure", "pneumonia")){
    stop('invalid outcome')
  } else if (is.numeric(rank)) {
    si <- which(fd[, "state"] == state)
    ts <- fd[si, ]                     # extracting dataframe for the called state
    ts[, eval(outcome)] <- as.numeric(ts[, eval(outcome)])
    ts <- ts[order(ts[, eval(outcome)], ts[, "hospital"]), ]
    output <- ts[, "hospital"][rank]
  } else if (!is.numeric(rank)){
    if (rank == "best") {
      output <- best(state, outcome)
    } else if (rank == "worst") {
      si <- which(fd[, "state"] == state)
      ts <- fd[si, ]    
      ts[, eval(outcome)] <- as.numeric(ts[, eval(outcome)])
      ts <- ts[order(ts[, eval(outcome)], ts[, "hospital"], decreasing = TRUE), ]
      output <- ts[, "hospital"][1]
    } else {
      stop('invalid rank')
    }
  }
  return(output)
}
# try out
rankhospital("TX", "heart failure", 4)
```

```{r}
rankhospital("NC", "heart attack", "worst")
```
```{r}
rankhospital("WA", "heart attack", 7)
```

```{r}
rankhospital("TX", "pneumonia", 10)
```

```{r}
rankhospital("NY", "heart attack", 7)
```

```{r}
rankall <- function(outcome, num = "best"){
  ## Read outcome data
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
  fd   <- as.data.frame(cbind(data[, 2],  # hospital
                              data[, 7],  # state
                              data[, 11],  # heart attack
                              data[, 17],  # heart failure
                              data[, 23]), # pneumonia
                        stringsAsFactors = FALSE)
  colnames(fd) <- c("hospital", "state", "heart attack", "heart failure", "pneumonia")
  fd[, eval(outcome)] <- as.numeric(fd[, eval(outcome)])
  
  ## Check that state and outcome are valid
  
  if (!outcome %in% c("heart attack", "heart failure", "pneumonia")){
    stop('invalid outcome')
  } else if (is.numeric(num)) {
    by_state <- with(fd, split(fd, state))
    ordered  <- list()
    for (i in seq_along(by_state)){
      by_state[[i]] <- by_state[[i]][order(by_state[[i]][, eval(outcome)], 
                                           by_state[[i]][, "hospital"]), ]
      ordered[[i]]  <- c(by_state[[i]][num, "hospital"], by_state[[i]][, "state"][1])
    }
    result <- do.call(rbind, ordered)
    output <- as.data.frame(result, row.names = result[, 2], stringsAsFactors = FALSE)
    names(output) <- c("hospital", "state")
  } else if (!is.numeric(num)) {
    if (num == "best") {
      by_state <- with(fd, split(fd, state))
      ordered  <- list()
      for (i in seq_along(by_state)){
        by_state[[i]] <- by_state[[i]][order(by_state[[i]][, eval(outcome)], 
                                             by_state[[i]][, "hospital"]), ]
        ordered[[i]]  <- c(by_state[[i]][1, c("hospital", "state")])
      }
      result <- do.call(rbind, ordered)
      output <- as.data.frame(result, stringsAsFactors = FALSE)
      rownames(output) <- output[, 2]
    } else if (num == "worst") {
      by_state <- with(fd, split(fd, state))
      ordered  <- list()
      for (i in seq_along(by_state)){
        by_state[[i]] <- by_state[[i]][order(by_state[[i]][, eval(outcome)], 
                                             by_state[[i]][, "hospital"], 
                                             decreasing = TRUE), ]
        ordered[[i]]  <- c(by_state[[i]][1, c("hospital", "state")])
      }
      result <- do.call(rbind, ordered)
      output <- as.data.frame(result, stringsAsFactors = FALSE)
      rownames(output) <- output[, 2]
    } else {
      stop('invalid num')
    }
  }
  return(output)
}
```

```{r}
r <- rankall("heart attack", 4)
as.character(subset(r, state == "HI")$hospital)
```
```{r}
r <- rankall("pneumonia", "worst")
as.character(subset(r, state == "NJ")$hospital)
```

```{r}
r <- rankall("heart failure", 10)
as.character(subset(r, state == "NV")$hospital)
```

