---
title: "R Notebook"
output: html_notebook
---

# Textual Data Formats

dumping and dputing are useful because the resulting textual format is edit-able, and in the case of corruption, potentially recoverable.
· Unlike writing out a table or csv file, dump and dput preserve the metadata (sacrificing some readability), so that another user doesn’t have to specify it all over again.
· Textual formats can work much better with version control programs like subversion or git which can only track changes meaningfully in text files
· Textual formats can be longer-lived; if there is corruption somewhere in the file, it can be easier to fix the problem
· Textual formats adhere to the “Unix philosophy”
· Downside: The format is not very space-efficient

## dput-ting R Objects

Another way to pass data around is by deparsing the R object with dput and reading it back in using dget.

```{r}
y <- data.frame(a = 1, b = "a")
dput(y)
```
```{r}
dput(y, file = "y.R")
```

```{r}
new.y <- dget("y.R")
new.y
```
dput single R object, dump multile R object

## Dumping R Objects


Multiple objects can be deparsed using the dump function and read back in using source.

```{r}
x<-"foo"
y <- data.frame(a = 1, b = "a")
dump(c("x", "y"), file = "data.R")
```

```{r}
rm(x, y)
```
```{r}
source("data.R")
y
```

