---
title: "R Notebook"
output: html_notebook
---

# Connections: Interfaces to the Outside World

Data are read in using connection interfaces. Connections can be made to files (most common) or to other more exotic things.

- file, opens a connection to a file    
- gzfile, opens a connection to a file compressed with gzip   
- bzfile, opens a connection to a file compressed with bzip2    
- url, opens a connection to a webpage    

## File Connections

```{r}
str(file)
```

description is the name of the file · open is a code indicating
- “r”readonly
- “w”writing(andinitializinganewfile)
- “a”appending
- “rb”,“wb”,“ab”reading,writing,orappendinginbinarymode(Windows)

## Connections

In general, connections are powerful tools that let you navigate files or other external objects. In practice, we often don’t need to deal with the connection interface directly.

```{r}
con <- file("foo.txt", "r") 
data <- read.csv(con) 
close(con)
```

is the same as

```{r}
data <- read.csv("foo.txt")
```

## Reading Lines of a Text File

```{r}
con <- gzfile("words.gz")
x <- readLines(con, 10)
x
```

```
>x
[1] "1080" "10-point" "10th" "11-point" [5] "12-point" "16-point" "18-point" "1st"
[9] "2" "20-point"
```


## Reading Lines of a Text File

readLines can be useful for reading in lines of webpages

```{r}
## This might take time
con <- url("http://www.jhsph.edu", "r")
x <- readLines(con)
head(x)
```


